Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de~Rham complex over convex domains and domains with shellable triangulations

### 2025-01-17 draft.tex

>>>>

\section{Previous geometric deformation estimate}

\begin{proposition}\label{proposition:starreflection}
    Let $\calT$ be a triangulation of an $n$-dimensional domain. 
    Let $S \in \calT$ be an inner simplex of dimension $\ell < n$,
    let $T \in \patch_{\calT}(S)$ be of dimension $n$,
    and let 
    \begin{align*}
        \Makoto := \overline{ \underlying{\patch_{\calT}(S)} \setminus T },
        \qquad 
        {\Gamma_1} := \Makoto \cap T,
        \qquad 
        {\Gamma_2} := \overline{ \partial T \setminus \partial \Makoto }.
    \end{align*}
    The following holds, where the constants on the right-hand sides are as stated in the proof. 
    \begin{enumerate}
    \item 
    There exists a bi-Lipschitz piecewise affine mapping
    \begin{align*}
        \Xi_{1} : T \rightarrow \Makoto
    \end{align*}
    which is the identity along ${\Gamma_1}$. 
    At any point, the singular values $\sigma_1 \geq \dots \geq \sigma_n$ of its Jacobian satisfy 
    \begin{gather*}
        \sigma_1 \leq \Cfive{n}{\ell}(\calT),
        \quad 
        \sigma_2,\dots,\sigma_n \leq \Cfiveprime{n}{\ell}(\calT),
        \quad 
        | \det\Jacobian\Xi_{1} |      \leq \Cdetfive{n}{\ell}(\calT),
        \\
        \sigma_n^{-1} \leq \Csechs{n}{\ell}(\calT),
        \quad 
        \sigma_1^{-1},\dots,\sigma_{n-1}^{-1} \leq \Csechsprime{n}{\ell}(\calT),
        \quad 
        | \det\Jacobian\Xi_{1}^{-1} | \leq \Cdetsechs{n}{\ell}(\calT).
    \end{gather*}
    Moreover, for any $0 \leq k \leq n$ and $p \in [1,\infty]$,
    \begin{align*}
        \sigma_{1}\cdots\sigma_{k} | \det\Jacobian\Xi_{1} |^{-\frac 1 p}
        \leq 
        \Cfivetrafo{n}{k}{\ell}{p}(\calT),
        \qquad 
        \sigma_{n}^{-1}\cdots\sigma_{n-k+1}^{-1} | \det\Jacobian\Xi_{1} |^{\frac 1 p}
        \leq 
        \Csechstrafo{n}{k}{\ell}{p}(\calT).
    \end{align*}

    \item 
    There exists a bi-Lipschitz piecewise affine mapping
    \begin{align*}
        \Xi_{2} : \underlying{\patch_{\calT}(S)} \rightarrow \Makoto
    \end{align*}
    which is the identity along $\Gamma_2$. 
    At any point, the singular values $\sigma_1 \geq \dots \geq \sigma_n$ of its Jacobian satisfy 
    \begin{gather*}
        \sigma_1 \leq \Csieb{n}{\ell}(\calT),
        \quad 
        \sigma_2,\dots,\sigma_n \leq \Csiebprime{n}{\ell}(\calT),
        \quad 
        | \det\Jacobian\Xi_{2} |      \leq \Cdetsieb{n}{\ell}(\calT),
        \\
        \sigma_n^{-1} \leq \Cacht{n}{\ell}(\calT),
        \quad 
        \sigma_1^{-1},\dots,\sigma_{n-1}^{-1} \leq \Cachtprime{n}{\ell}(\calT),
        \quad 
        | \det\Jacobian\Xi_{2}^{-1} | \leq \Cdetacht{n}{\ell}(\calT).
    \end{gather*}
    Moreover, for any $0 \leq k \leq n$ and $p \in [1,\infty]$,
    \begin{align*}
        \sigma_{1}\cdots\sigma_{k} | \det\Jacobian\Xi_{2} |^{-\frac 1 p}
        \leq 
        \Csiebtrafo{n}{k}{\ell}{p}(\calT),
        \qquad 
        \sigma_{n}^{-1}\cdots\sigma_{n-k+1}^{-1} | \det\Jacobian\Xi_{2} |^{\frac 1 p}
        \leq 
        \Cachttrafo{n}{k}{\ell}{p}(\calT).
    \end{align*}
    % DONE {TCF: Should $\sigma_n$ be $\sigma_n^{-1}$ in the second line of both statements?}
    \end{enumerate}
\end{proposition}
\begin{proof}
    We derive the estimate in several steps. 
    In what follows, we use the notation $\hat z$ for the normalization of any vector $z \in \bbR^{n}$.
    \begin{itemize}
        \item 
        Without loss of generality, after shifting the coordinate system, the barycenter $z_{S}$ of $S$ is the origin. 
        % Let $I := \Makoto \cap T$, which is the union of $\ell$ different faces of $T$, with $1 \leq \ell \leq n$.
        %We let ${\Gamma_2} \subseteq \partial T$ be the union of the remaining $n+1-\ell$ faces of $T$.
        %So ${\Gamma_2$} is within the boundary of the star $|\patch_{\calT}(S)|$ and is disjoint from $S$.
        We fix the subsimplex $S' \subseteq T$ that is complementary to $S$ within the simplex $T$. 
        In other words, the vertices of $S'$ are the vertices of $T$ that are not vertices of $S$. 
        Note that ${\Gamma_2} \subseteq \partial T$ is the union of exactly those faces of $T$ that contain $S'$,
        whereas ${\Gamma_1} = \Makoto \cap T$ is the union of exactly those faces of $T$ that contain $S$.

        \item 
        We let $z_{S'}$ be the midpoint of $S'$. 
        We apply barycentric refinement to the complementary simplex $S'$, which produces a new triangulation $\calD$ of the underlying set $\underlying{T}$. 
        Obviously, all $n$-simplices in $\calD$ contain $S$ and $z_{S'}$.

        We define another simplicial complex $\calD^c$ as follows:
        given any $n$-simplex $K \in \calD$, 
        we replace its vertex $z_{S}'$ by the new vertex $-z_{S}'$ at the opposite position,
        thus obtaining a new simplex $K^{c}$. 
        Indeed, that $\calD^c$ is a simplicial complex follows easily from $\calD$ being a simplicial complex.

        We let the simplicial complex $\calD^\ast$ be the union of $\calD$ and $\calD^c$. 
        By construction, all its $n$-simplices contain $S$ as a subsimplex,
        % Indeed, $\calD^\ast$ is indeed a triangulation since they are disjoint except for their proper subsimplices. 
        Moreover, $S$ is an inner subsimplex of that triangulation. 
        In particular, $\calD^\ast$ is its own star around $S$.

        % TODO Check that $\calD^\ast$ is a triangulation

        \item 
        As additional step, we introduce another simplicial complex $\calG^\ast$ from $\calD^\ast$ via barycentric subdivision of $S$.
        Then $\calG^\ast$ is its own local star around the barycenter $z_S$.
        In particular, all $n$-simplices in $\calG^\ast$ contain $z_{S}$. 
        The simplicial complex $\calG^\ast$ also contains as subcomplexes the corresponding refinements of $\calD$ and $\calD^{c}$, 
        which we denote by $\calG$ and $\calG^{c}$, respectively.
        Given any $n$-simplex in $\calG$, replacing its vertex $z_{S}$ with its opposite $-z_{S}$ gives an associated $n$-simplex in $\calG^{c}$.

        Lastly, we introduce one more simplicial complex $\calK$ obtained from $\patch_{\calT}(S)$ via barycentric refinement of $S$.
        % Notice that $A_{G^{c}}$ and $A_{G}$ have the same singular values. Thus
        %     \matnorm{ \Jacobian \Theta_{|G}      } &\leq \| A_{G^{c}} \| \cdot \| A_{G}^{-1} \| \leq \geometricshapemeasure(A_{G}),
        %     \matnorm{ \Jacobian \Theta_{|G}^{-1} } &\leq \| A_{G^{c}}^{-1} \| \cdot \| A_{G} \| \leq \geometricshapemeasure(A_{G}).
        We introduce a new mapping $\Theta : \underlying{\calG} \rightarrow \underlying{\calG^c}$ between the underlying sets as follows.
        Let $G \in \calG$ be an $n$-simplex and $G^{c} \in \calG^{c}$ be constructed from $G$.
        We let $A_{G} : \Delta^{n} \rightarrow G$ and $A_{G^{c}} : \Delta^{n} \rightarrow G^{c}$
        be affine reference transformations
        that agree on the vertices common to $G$ and $G^{c}$.
            \Theta_{|G} := A_{G^{c}} \circ A_{G}^{-1}
        preserves volumes:
            |\det( \Jacobian      \Theta_{|G} )|
            |\det( \Jacobian^{-1} \Theta_{|G} )|
        We want to characterize the singular values of this transform. 
        Let $h_{z} \in \bbR^{n}$ be the height vector of $z_{S'}$ inside the simplex $G$.
        We verify that 
            \Theta_{|G}(x) 
            2 \frac{\langle \hat h_{z}, x \rangle}{\langle \hat h_{z}, \hat z_{S'} \rangle}
            \hat z_{S'}
        Indeed, we check that the right-hand side equals $x$ whenever $x$ lies in the plane orthogonal to $h_{z}$,
        and that it equals $-z_{S'}$ when $x = z_{S'}$.
%         and when $x = z_{S'}$, then 
%             -z_{S'} 
%             =
%             z_{S'}
%             - 
%             2 \hat z_{S'} \frac{\langle \hat h_{z}, z_{S'} \rangle}{\langle \hat h_{z}, \hat z_{S'} \rangle}
        If we orthogonally decompose $z_{S'} = h_{z} + b_{z}$ for some $b_{z} \in \bbR^{n}$, then 
            \Theta_{|G}( h_{z} ) 
            h_{z}
            2 \frac{\langle h_{z}, h_{z} \rangle}{\langle h_{z}, z_{S'} \rangle} z_{S'}
            h_{z}
            2 \frac{\langle h_{z}, z_{S'} \rangle}{\langle h_{z}, z_{S'} \rangle} z_{S'}
            %\\&
            h_{z}
            2 z_{S'}
            %\\&
            - h_{z}
            2 b_{z}
        Evidently, the transformation $\Theta_{|G}$ equals the identity on the orthogonal complement of the span of $h_{z}$ and $b_{z}$. 
        Let $\beta$ be the angle between $z_{S'}$ and $h_{z}$. 
        Then $\vecnorm{ h_{z} } = \cos(\beta) \vecnorm{ z_{S'} }$ and $\vecnorm{ b_{z} } = \sin(\beta) \vecnorm{ z_{S'} }$. 
        We study the singular values of the matrix 
            M_{\Theta,G} 
            := 
            \begin{pmatrix}
            -1                 & 0
            -2 \vecnorm{ b_{z} }/\vecnorm{ h_{z} } & 1
            \end{pmatrix}
            \begin{pmatrix}
            -1             & 0
            -2 \tan(\beta) & 1
            \end{pmatrix}
        We compute the eigenvalues of the symmetric matrix $M_{\Theta,G}^{\ast} M_{\Theta,G}$. 
        Its square roots are the desired singular values of $M_{\Theta,G}$:
            \sigma_{\max}(\Theta,G) := \sqrt{ 2\tan(\beta)^2 + 1 + 2 \tan(\beta) \sqrt{ \tan(\beta)^2 + 1 } } = \sqrt{ 1 + \tan^{2}(\beta) } + \tan(\beta) \label{math:theta_max}
            \sigma_{\min}(\Theta,G) := \sqrt{ 2\tan(\beta)^2 + 1 - 2 \tan(\beta) \sqrt{ \tan(\beta)^2 + 1 } } = \sqrt{ 1 + \tan^{2}(\beta) } - \tan(\beta) \label{math:theta_min}
        They are mutual reciprocals of each other, and they are monotonically increasing and decreasing, respectively, in $\tan(\beta)$.
        We conclude that these two are also the maximal and minimal singular values of $\Jacobian\Theta_{|G}$,
        whereas all its other singular values equal $1$.

%         Let $\beta$ be the angle between $z_{S'}$ and $h_{z}$. 
        We develop explicit bounds for these singular values. 
        The definition of the tangent and the decomposition $z_{S'} = h_{z} + b_{z}$ 
        imply that $\tan(\beta) = \vecnorm{ b_{z} } / \vecnorm{ h_{z} }$.
        % Obviously, $\vecnorm{ b_{z} } \leq \vecnorm{ z_{S'} }$.

        Recall that $G \in \calG$ is obtained from $T$ via barycentric subdivisions, first at $z_{S}$ and then at $z_{S'}$. 
        Let $F_{z} \subseteq G$ be the face opposite to the vertex $z_{S'}$,
        which must be contained in some face of $T$. 
        Since $S'$ has dimension $n-\ell-1$, the vector $(n-\ell) h_{z}$ is the height vector of that face of $T$. 
        Thus, 
        \begin{align*}
            \frac{ \vecnorm{ b_{z} } }{ \vecnorm{ h_{z} } }
            = 
            \frac{ \vecnorm{ b_{z} } }{ (n-\ell)^{-1} \vecnorm{ (n-\ell) h_{z} } }
            = 
            \frac{ \vecnorm{ z_{S'} } }{ (n-\ell)^{-1} \vecnorm{ (n-\ell) h_{z} } }
            \leq 
            (n-\ell) \aspectratio(T)
%             = 
%             \frac{ \vecnorm{ b_{z} } \vol(F_{z}) }{ (n-\ell)^{-1} n \vol(T) }
%             \leq 
%             \frac{ n-\ell }{n!}
%             \cdot 
%             \frac{ \diam(T)^{n} }{ \vol(T) }
            .
        \end{align*}
        This establishes bounds on the singular values of the transformation. 
        We abbreviate 
        \begin{align}\label{math:upper_bound_for_theta}
            \mu_{T,\ell} 
            := 
%             \sqrt{ 1 + \tan^{2}(\beta) } - \tan(\beta) 
%             \geq 
            \sqrt{ 
                1 + (n-\ell)^{2} \aspectratio(T)^{2} % \frac{ \diam(T)^{2n} }{ n!^{2}\vol(T)^{2} } 
            } 
            + 
            (n-\ell) \aspectratio(T) % \frac{ \diam(T)^{n} }{ n! \vol(T) }            
            .
        \end{align}
        In summary, the singular values of the Jacobian of $\Theta$ at almost every $x$ satisfy 
        \begin{align}
            \sigma_{1}(\Theta,x) \leq \mu_{T,\ell},
            \quad 
            \sigma_{2}(\Theta,x) = \dots = \sigma_{n-1}(\Theta) = 1,
            \quad 
            \sigma_{n}(\Theta,x)^{-1} \leq \mu_{T,\ell}.
        \end{align}
        \color{red}It is also evident that the eigenvalues of the Jacobian almost everywhere have magnitude $1$.\color{black}

        \item 
        We introduce another mapping $\Phi : \underlying{\calG^\ast} \rightarrow \underlying{\calG^{c}}$ as follows. 
        Consider any $n$-simplex $G \in \calG$ and let $G^c \in \calG^{c}$ be its image under $\Theta$.
        We construct a bi-Lipschitz mapping 
        \begin{align*}
            \Phi_{G} : G \cup G^{c} \rightarrow G^{c}.
        \end{align*}
        The construction will be such that the union of $\Phi_{G}$ for all $n$-simplices $G \in \calG$
        will define the desired bi-Lipschitz mapping $\Phi : \underlying{\calG^\ast} \rightarrow \underlying{\calG^{c}}$,
        which will be the identity along $\partial\underlying{\calG^\ast} \setminus \partial\underlying{\calG}$.

        Once again, $h_{z}$ denotes the height of $z_{S'}$ within $G$.
        Here, we let $Q \subseteq G \cap G^{c}$ be the subsimplex 
        that is complementary to the line segment from the origin to $z_{S'}$ in $G$.
        Equivalently, $Q$ is complementary to the line segment from the origin to $-z_{S'}$ in $G^{c}$.
        From the definition of simplices, we now conclude that any $x \in G \cup G^{c}$ 
        has a unique representation
        \begin{align*}
            x = \lambda z_{S'} + \mu x_{Q}, \quad \lambda \in [-1,1], \quad \mu \in [0,1], \quad |\lambda| + \mu \leq 1, \quad x_{Q} \in Q,
        \end{align*}
        Since $\mu x_{Q}$ lies in the hyperplane spanned by the origin and $Q$, we have 
        \begin{align*}
            \lambda = \lambda(x) := \frac{\langle h_{z},x\rangle}{\langle h_{z},z_{S'}\rangle} 
            .
        \end{align*}
        Based on that observation, we define 
        \begin{align*}
            \Phi_{G}(x) % TODO : Sign error. Check values for 0 and 1. Then proceed from there ... 
            &
            := 
            \mu x_{Q} + \frac 1 2 \left( \lambda(x) - 1 \right) z_{S'}
            \\&
            = 
            x - \lambda(x) z_{S'} + \frac{ \lambda(x)}{ 2 } z_{S'} - \frac{ 1 }{ 2 } z_{S'}
            \\&
            = 
            x - \frac{ \lambda(x) }{2} z_{S'} - \frac{ 1 }{ 2 } z_{S'}
            \\&
            = 
            x - \frac{\langle h_{z},x\rangle}{2\langle h_{z},z_{S'}\rangle} z_{S'} - \frac 1 2 z_{S'}
            .
        \end{align*}
        We readily verify that this transformation is a bi-Lipschitz mapping from $G \cup G^{c}$ onto $G^{c}$
        that satisfies the desired mapping properties. 
        It remains to analyze its Jacobian to get explicit estimates.

        We once more introduce an orthogonal decomposition $h_{z} + b_{z} = z_{S'}$ for some $b_{z} \in \bbR^{n}$.
        With that,
            \Jacobian \Phi_{G}(x)
            \frac{1}{ 2 \langle h_{z},z_{S'}\rangle }  
            z_{S'} \otimes h_{z}^{t}
            \frac{1}{ 2 \langle h_{z},z_{S'}\rangle } 
            h_{z} \otimes h_{z}^{t}
            \frac{1}{ 2 \langle h_{z},z_{S'}\rangle } 
            b_{z} \otimes h_{z}^{t}
            \frac{1}{ 2 \langle h_{z}, h_{z}\rangle } 
            h_{z} \otimes h_{z}^{t}
            \frac{1}{ 2 \langle h_{z}, h_{z}\rangle } 
            b_{z} \otimes h_{z}^{t}
            \frac{1}{ 2 } 
            \hat h_{z} \otimes \hat h_{z}^{t}
            \frac{ 1 \vecnorm{ b_{z} } }{ 2 \vecnorm{ h_{z} } } 
            \hat b_{z} \otimes \hat h_{z}^{t}
        The Jacobian acts as the identity over the orthogonal complement of the span of $h_{z}$ and $z_{s'}$.
        We write $\beta$ for the angle between $h_{z}$ and $z_{S'}$.
        Hence, $\tan(\beta) = \vecnorm{ b_{z} } / \vecnorm{ h_{z} }$. 
        We verify that 
            \Jacobian \Phi_{G}(x) \cdot \hat b_{z} = \hat b_{z},
            \Jacobian \Phi_{G}(x) \cdot \hat h_{z} = \frac 1 2 h_{z} - \frac{ 1 \vecnorm{ b_{z} } }{ 2 \vecnorm{ h_{z} } } \hat b_{z}
        It remains to study the singular values of the matrix 
            M_{\Phi,G} 
            \begin{pmatrix}
                 \frac 1 2             & 0 
                -\frac 1 2 \tan(\beta) & 1 
            \end{pmatrix}
%             = 
%             \frac 1 2 
%             \begin{pmatrix}
%                 -1 & 0 
%                 -3\tan(\beta) & 2 
%             \end{pmatrix}
        By building the symmetric matrix $M_{\Phi,G}^{t} M_{\Phi,G}$ and computing its eigenvalues, 
        we obtain the singular values 
            \sigma_{\max}(\Phi,G) 
            \frac{1}{4} \sqrt{ 9 + \tan(\beta)^2 } + \frac{1}{4} \sqrt{ 1 + \tan(\beta)^2 } 
            \sigma_{\min}(\Phi,G) 
            \frac{1}{4} \sqrt{ 9 + \tan(\beta)^2 } - \frac{1}{4} \sqrt{ 1 + \tan(\beta)^2 } 
        These are monotonically increasing from $1$ and decreasing from $0.5$, respectively, in $\tan(\beta)$. 
        Hence, these are also the maximal and minimal singular values of the Jacobian $\Jacobian \Phi_{G}$, the remaining singular values being equal to $1$. 
        Notice that $\sigma_{\max}(\Phi,G) \sigma_{\min}(\Phi,G) = 1/2$ and 
            \det \Jacobian\Phi = \frac 1 2.
        We now recall that the height of $h_{z}$ in $G \in \calG$ equals $(\ell+1)^{-1}$ multiplied with the height of some vertex of $S$ within $T$.
        Similar as above, we use the upper bound 
            \tan(\beta) = \frac{\vecnorm{ b_{z} }}{\vecnorm{ h_{z} }} \leq (\ell+1) \aspectratio(T).
        We conclude that the singular values of the Jacobian of $\Phi$ at almost every $x$ satisfy 
        \begin{gather}
            \sigma_{1}(\Phi,x) % \matnorm{ \Jacobian\Phi_{G} } 
            \frac 1 4 \sqrt{ 9 + (\ell+1)^2 \aspectratio(T)^2 } + \frac 1 4 \sqrt{ 1 + (\ell+1)^2 \aspectratio(T)^2 }
            1 + \frac{1}{2} (\ell+1) \aspectratio(T) 
            \label{math:phi_max}
            \sigma_{2}(\Phi,x) = \dots = \sigma_{n-1}(\Phi,x) = 1,
            \sigma_{n}(\Phi,x)^{-1} % \matnorm{ \Jacobian\Phi_{G}^{-1} } 
            \frac 1 2 \sqrt{ 9 + (\ell+1)^2 \aspectratio(T)^2 } + \frac 1 2 \sqrt{ 1 + (\ell+1)^2 \aspectratio(T)^2 }
            2 + (\ell+1) \aspectratio(T) 
            \label{math:phi_min}
        \end{gather}
        \color{red}Moreover, it is evident that the eigenvalues of the Jacobian are $1$ with multiplicity $n-1$ and $\tfrac 1 2$ with multiplicity $1$.\color{black}
        This finishes the discussion of the transformation $\Phi$.

        \item 
        We continue with an auxiliary result.
        Let $x, h_1, h_2 \in \bbR^{n}$ be non-zero. 
        Define 
        \begin{align*}
            F(x) 
            = 
            \frac{ \vecnorm{h_2}^{2} }{ \langle h_2, x \rangle }
            \frac{ \langle h_1, x \rangle }{ \vecnorm{ h_1 }^{2} }
            x
            .
        \end{align*}
        This is defined away from the hyperplane orthogonal to $h_2$.
        If $x \in \bbR^{n}$ with $\langle h_1, x \rangle = \vecnorm{ h_1 }^{2}$, then $\langle F(x), h_2 \rangle = \vecnorm{ h_2 }^{2}$.
        That shows that $F(x)$ is a radial mapping that maps each point on the hyperplane defined by the normal vector $h_1$
        onto the colinear point that lies on the hyperplane defined by the normal vector $h_2$. Indeed,

        We identify the Lipschitz properties of this mapping by computing its Jacobian. 
        We write $\alpha_1, \alpha_2 \geq 0$ for the two angles between $h_1$ and $x$ and between $h_2$ and $x$, respectively. 
        In what follows, $y \in \bbR^{n}$.
        \begin{align*}
            \Jacobian F(x) \cdot y
            &=
            \frac{ \vecnorm{ h_2 }^{2} }{ \langle h_2, x \rangle }
            \frac{ \langle h_1, y \rangle }{ \vecnorm{ h_1 }^{2} }
            x
            - 
            \frac{ \langle h_2, h_2 \rangle \langle h_2, y \rangle }{ \langle h_2, x \rangle^{2} }
            \frac{ \langle x, h_1 \rangle }{ \langle h_1, h_1 \rangle }
            x 
            +
            \frac{ \langle h_2, h_2 \rangle }{ \langle h_2, x \rangle }
            \frac{ \langle x, h_1 \rangle }{ \langle h_1, h_1 \rangle }
            y 
            \\&=
            \frac{ \langle h_2, h_2 \rangle }{ \vecnorm{ h_2 } \vecnorm{x} \cos(\alpha_2) }
            \left( y - \frac{ \langle h_2, y \rangle }{ \langle h_2, x \rangle } x \right)
            \frac{ \langle x, h_1 \rangle }{ \langle h_1, h_1 \rangle }
            +
            \frac{ \vecnorm{ h_2 }^{2} }{ \langle h_2, x \rangle }
            \frac{ \langle h_1, y \rangle }{ \vecnorm{ h_1 }^{2} }
            x
            \\&=
            \frac{ \langle \hat h_2, h_2 \rangle }{ \cos(\alpha_2) }
            \left( y - \frac{ \langle \hat h_2, y \rangle }{ \langle \hat h_2, \hat x \rangle } \hat x \right)
            \frac{ \langle \hat x, \hat h_1 \rangle }{ \langle h_1, \hat h_1 \rangle }
            +
            \frac{ \vecnorm{ h_2 } }{ \langle \hat h_2, \hat x \rangle }
            \frac{ \langle \hat h_1, y \rangle }{ \vecnorm{ h_1 } }
            \hat x
            \\&=
            \frac{ \vecnorm{ h_2 } }{ \cos(\alpha_2) }
            \left( y - \frac{ \langle \hat h_2, y \rangle }{ \cos(\alpha_2) } \hat x \right)
            \frac{ \cos(\alpha_1) }{ \vecnorm{ h_1 } }
            +
            \frac{ \vecnorm{ h_2 } }{ \cos(\alpha_2) }
            \frac{ \langle \hat h_1, y \rangle }{ \vecnorm{ h_1 } }
            \hat x
            \\&=
            \frac{ \vecnorm{ h_2 } }{ \vecnorm{ h_1 } }
            \frac{ \cos(\alpha_1) }{ \cos(\alpha_2) }
            \left( 
                y 
                - 
                \frac{ \langle \hat h_2, y \rangle }{ \cos(\alpha_2) } \hat x 
                + 
                \frac{ \langle \hat h_1, y \rangle }{ \cos(\alpha_1) } \hat x 
            \right)
            .
        \end{align*}
        We define $a := \cos(\alpha_1)^{-1} \hat h_1 - \cos(\alpha_2)^{-1} \hat h_2$. 
        Notice that $a \perp \hat x$. The matrix $\Id + \hat x \otimes a^{t}$ is the identity over the orthogonal complement of the linear span of $a$ and $x$. Over that two-dimensional subspace, it can be represented by a triangular matrix.
        Similar as above, its largest and the smallest singular values $\sigma_{max} \geq 1 \geq \sigma_{\min}$ are
        %~\cite{blinn1996consider} 
        \begin{align*}
            \sigma_{\max} = \sqrt{ 1 + \frac{ \vecnorm{a}^{2} }{ 4 } } + \frac{ \vecnorm{a} }{2},
            \qquad 
            \sigma_{\min} = \sqrt{ 1 + \frac{ \vecnorm{a}^{2} }{ 4 } } - \frac{ \vecnorm{a} }{2}.     
        \end{align*}
        Notice that these are strictly monotonically increasing or decreasing, respectively, in $\vecnorm{a}$.
%         We thus consider the upper bound 
%         \begin{align*}
%             \vecnorm{a} 
%             \leq 
%             2 \max_{1 \leq i \leq 2} 
%             \left( 
%                 \frac{ \hat h_i }{ \langle \hat x, \hat h_i \rangle } 
%             \right)
%             .
%         \end{align*}
        \color{red}Lastly, the matrix $\Id + \hat x \otimes a^{t}$ has the eigenvalue $1$ with multiplicity $n-1$ and a trace of $n + \langle a, \hat x \rangle$,
        so it has one eigenvalue $1 + \langle a, \hat x \rangle$.\color{black}
        We now describe the construction of a bi-Lipschitz mapping 
            \Psi : \underlying{\calG^\ast} \rightarrow \underlying{\calK}.
        Recall that $\calG^\ast$ and $\calK$ are their own respective stars around the origin,
        so they must contain an open ball around the origin. 
        % TODO \mwl{Here we need to show that it is indeed a ball. It must be a domain triangulation.}
        Let ${G} \in \calG^\ast$ be an $n$-dimensional simplex and let $x \in {G}$ be non-zero. There exists a simplex $K \in \calK$ that intersects the ray $\bbR_0^+ \cdot x$. We let $h_{{G}}$ and $h_{K}$ be the normal vectors of the origin within ${G}$ and $K$, respectively. They define respective hyperplanes $H_{G}$ and $H_{K}$,
        and we write $F_{G} = G \cap H_{G}$ and $F_{K} = K \cap H_{K}$ for the corresponding faces. 
        We notice that $x$ has a sharp angle to both $h_{{G}}$ and $h_{K}$,
        that is, the scalar product of $x$ with either height vector is positive. 
            \frac{ \vecnorm{ h_K }^{2} }{ \langle h_K, x \rangle }
            \frac{ \langle h_{{G}}, x \rangle }{ \vecnorm{ h_{{G}} }^{2} }
        By the same line of thought, we see that it is invertible with inverse satisfying 
            \Psi^{-1}(z) 
            \frac{ \vecnorm{ h_{{G}} }^{2} }{ \langle h_{{G}}, z \rangle }
            \frac{ \langle h_K, z \rangle }{ \vecnorm{ h_K }^{2} }
            z,
            z = \Psi(x)
        We let $\alpha_K$ and $\alpha_{{G}}$ be the angles of $x$ with the vectors $h_{{G}}$ and $h_{K}$, respectively,
        and let 
            a := \cos(\alpha_{{G}})^{-1} \hat h_{{G}} - \cos(\alpha_K)^{-1} \hat h_K
        The preceding auxiliary results now contribute all information on the singular values of these Jacobians:
            % \matnorm{ \Jacobian\Psi(x) }
            \sigma_{1}( \Jacobian\Psi(x) )
            \frac{ \vecnorm{ h_K } }{ \vecnorm{ h_{{G}} } }
            \frac{ \cos(\alpha_{{G}}) }{ \cos(\alpha_K) }
                \sqrt{ 1 + \frac{ \vecnorm{a}^{2} }{ 4 } } + \frac{ \vecnorm{a} }{2}
            % \matnorm{ \Jacobian\Psi(z)^{-1} }
            \sigma_{1}( \Jacobian\Psi(z)^{-1} )
            \frac{ \vecnorm{ h_{{G}} } }{ \vecnorm{ h_K } }
            \frac{ \cos(\alpha_K) }{ \cos(\alpha_{{G}}) }
                \sqrt{ 1 + \frac{ \vecnorm{a}^{2} }{ 4 } } - \frac{ \vecnorm{a} }{2}
            \right)^{-1},
            \sigma_{2}( \Jacobian\Psi(x) ) = \dots = \sigma_{n-1}( \Jacobian\Psi(x) )
            \frac{ \vecnorm{ h_K } }{ \vecnorm{ h_{{G}} } }
            \frac{ \cos(\alpha_{{G}}) }{ \cos(\alpha_K) }
            \det\Jacobian\Psi(x)
            \frac{ \vecnorm{ h_K } }{ \vecnorm{ h_{{G}} } }
            \frac{ \cos(\alpha_{{G}}) }{ \cos(\alpha_K) }
            \right)^{n},
            \det\Jacobian\Psi^{-1}(z)
            \frac{ \vecnorm{ h_{{G}} } }{ \vecnorm{ h_{K} } }
            \frac{ \cos(\alpha_{K}) }{ \cos(\alpha_{{G}}) }
            \right)^{n}
        We let $x_{G} \in F_{G}$ and $x_{K} \in F_{K}$ be the intersection points of the ray $\bbR \cdot x$ with the respective faces.
        By the definition of the cosine, 
            \frac{ \vecnorm{ h_K } }{ \vecnorm{ h_{{G}} } }
            \frac{ \cos(\alpha_{{G}}) }{ \cos(\alpha_K) }
            \frac{ \vecnorm{ x_K } }{ \vecnorm{ x_{{G}} } }
            \frac{ \vecnorm{ x_K } }{ \vecnorm{ h_{{G}} } }
            \frac{ \diam(K) }{ \vecnorm{ h_{{G}} } }
            \frac{ \vecnorm{ h_{{G}} } }{ \vecnorm{ h_K } }
            \frac{ \cos(\alpha_K) }{ \cos(\alpha_{{G}}) }
            \frac{ \vecnorm{ x_{G} } }{ \vecnorm{ x_{K} } }
            \frac{ \vecnorm{ x_{G} } }{ \vecnorm{ h_{K} } }
            \frac{ \diam(G) }{ \vecnorm{ h_{K} } }
%             \frac{ \vecnorm{ h_K } }{ \vecnorm{ h_{{G}} } }
%             \frac{ \cos(\alpha_{{G}}) }{ \cos(\alpha_K) }
%             \frac{ \vecnorm{ x_K } }{ \vecnorm{ x_{{G}} } }
        We need further details to complete our estimates. 
        We know that $\vecnorm{ x_K } \leq \diam(K)$ and that the diameters of the simplices in $\calK$ are already bounded by the maximum diameter in $\patch_{\calT}(S)$. The height $h_{K}$ is obtained from a height vector of a vertex in $\patch_{\calT}(S)$ by scaling with the factor $(\ell+1)^{-1}$.
        That estimates the quotient $\vecnorm{ x_K } / \vecnorm{ h_K }$. 
        In order to estimate $\vecnorm{ h_{G} }$, we first notice the case $G \in \calG$ implies $\vecnorm{ h_{G} } \leq \diam(G) \leq \diam(T)$. 
        In the case $G \in \calG^{c}$, we recall that $x = \lambda (-z_{S'}) + (1-\lambda) z_{x}$ for some $\lambda \in [0,1]$ and $z_{x} \in G$
        that must lie in the face of $G$ that is opposite to $-z_{S'}$. 
        The preimage of $x$ under $\Theta_{{G}}$ is $\lambda z_{S'} + (1-\lambda) z_{x}$. 
        It follows again that $\vecnorm{ h_{G} } \leq \diam(T)$. 
        It remains to find a lower bound for $\vecnorm{ h_{G} }$.
        Every height of the origin in some simplex of $\calG^\ast$ is obtained from a height vector of some vertex in $S$ in some simplex in $\calD^\ast$
        by scaling with $(\ell+1)^{-1}$. If $G \in \calG$, then that original height was already a height of a vertex of $S$ in $T$. 
        If instead $G \notin \calG$, then each of the latter height vectors has length bounded from below by the height of some vertex in $S$ within a simplex of $\calD$, multiplied by $\mu_{T,\ell}$. The height of any vertex of $S$ in some simplex of $\calD$ is the same as the height $h_T$ of that vertex in the original simplex $T$. 
        %         In oder to estimate $\vecnorm{ x_{G} }$, recall that $x = \lambda (-z_{S'}) + (1-\lambda) z_{x}$ for some $\lambda \in [0,1]$ and $z_{x} \in G$
        %         being in the face of $G$ that is opposite to $z_{x}$. The preimage of $x$ under $\Theta_{{G}}$ is $\lambda z_{S'} + (1-\lambda) z_{x}$. 
        %         It follows that $\vecnorm{ x_{G} } \leq \diam(T)$. 
%         We know that whenever $T$ is a simplex with some face $F$ and $h_F$ is corresponding height vector, then 
%             \frac{\diam(T)}{ \vecnorm{ h_{F} }}
%             \aspectratio(T)
%             % \frac{ \vol(F) \diam(T) }{ n \vol(T) } \leq \frac{ \diam(T)^{n} }{ n! \vol(T) }
        Putting all these technical observations together, we estimate 
            \frac{ \vecnorm{ x_{K} } }{ \vecnorm{ h_{{G}} } }
            (\ell+1)\mu_{T,\ell}
            \frac{ \diam(K) }{ \vecnorm{ h_{T} } }
            (\ell+1)\mu_{T,\ell}
            \diameterratio(\calT)
            \frac{ \diam(T) }{ \vecnorm{ h_{T} } }
            (\ell+1)
            \diameterratio(\calT)
            \aspectratio(T) % \frac{ \diam(T)^{n} }{ n! \vol(T) }
            \mu_{T,\ell}
            \frac{ \vecnorm{ x_{G} } }{ \vecnorm{ h_{K} } }
            (\ell+1)
            \sup_{ K \in \patch_{\calT}(S) }
            \frac{ \diam(T) }{ \vecnorm{ h_{K} } }
            (\ell+1)
